--- 
title: |
  Text Based Survival Statistics
format:
  html:
    css: styles.css
---



```{r}
#| echo: false
#| warning: false
#| include: false
#| label: library_data_imports

library(tidyverse)
library(ggplot2)
library(phyloseq)
library(microViz)
library(survival)
library(survminer)
library(viridis)
library(gtsummary)
library(yingtools2)
source("utility_funcs.R")


```

This notebok is just to ensure statistics reported in the text are consistent and to make it clear wat set of variables was used in each.  Each section contains the line from the text followed by a section with the corresponding code. 

```{r}
#| warning: false
#| echo: false
#| label: load_data

gu_cohort_data <- readRDS("../data/gu_survey_cohort.RDS")
mixed_solid_tumor <- readRDS("../data/mixed_solid_tumor.RDS")
external_melanoma <- readRDS("../data/external_melanoma_metadata.RDS")
external_rcc <- readRDS("../data/external_rcc.rds")
```

Higher dietary fiber intake was associated with longer progression free survival (PFS) and overall survival (OS) both in analyses examining fiber intake as a continuous variable (p = 0.045 PFS, p = 0.048 OS) and as a categorical variable.
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: fiber_univariate

#line 167-170

cox_pfs_gu_uv_continous <-coxph(Surv(pfs_mo,
                               pod_status)~  aofib,
                          gu_cohort_data)

variable = "aofib"

sum_pfs_gu_uv_continous <- summary(cox_pfs_gu_uv_continous)
cat(paste(
  "-----------------------------------------------\n\np value unadjusted fiber as a continus variable, PFS:",
  round(sum_pfs_gu_uv_continous$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_gu_uv_continous$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_gu_uv_continous$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_gu_uv_continous$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

cox_os_gu_uv_continous <-coxph(Surv(os_mo,
                               os_status)~  aofib,
                          gu_cohort_data)

sum_os_gu_uv_continous <- summary(cox_os_gu_uv_continous)
cat(paste(
  "-----------------------------------------------\n\np value unadjusted fiber as a continus variable, OS:",
  round(sum_os_gu_uv_continous$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_os_gu_uv_continous$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_os_gu_uv_continous$conf.int[variable, "lower .95"],digits = 2),
  round(sum_os_gu_uv_continous$conf.int[variable, "upper .95"],digits = 3),
  "]"
))
```

Each 5-gram increase in daily fiber intake was associated with a 15% reduction in the risk of disease progression or death (hazard ratio [HR] of 0.85, 95% confidence interval [CI], 0.72 – 0.997, p = 0.045).
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: fiber_5g_univariate_pfs

#line 170-172
cox_pfs_gu_uv_5g <-coxph(Surv(pfs_mo,
                               pod_status)~  fiber_5g_increment,
                          gu_cohort_data)

variable = "fiber_5g_increment"
sum_pfs_gu_5g_uv <- summary(cox_pfs_gu_uv_5g)
cat(paste(
  "-----------------------------------------------\n\np value fiber in 5g increments, PFS:",
  round(sum_pfs_gu_5g_uv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_gu_5g_uv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_gu_5g_uv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_gu_5g_uv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

```
Compared to patients in the lowest tertile of fiber intake, patients in the highest tertile had a reduced risk of disease progression or death in both a univariable model (HR 0.43, 95% 0.22-0.83, p = 0.012) ...
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: fiber_tertile_univariate_pfs

#line 172-174
cox_pfs_gu_uv_tertile <-coxph(Surv(pfs_mo,
                               pod_status)~  fiber_tertile_factor,
                          gu_cohort_data)

variable = "fiber_tertile_factorT3 (highest)"
sum_pfs_gu_tertile_uv <- summary(cox_pfs_gu_uv_tertile)
cat(paste(
  "-----------------------------------------------\n\np value fiber highest vs lowest tertile univariate, PFS:",
  round(sum_pfs_gu_tertile_uv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_gu_tertile_uv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_gu_tertile_uv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_gu_tertile_uv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

```
... and a multivariable model adjusted for diagnosis, stage, and performance status, and stratified by age (HR 0.49, 95% CI 0.25-0.97, p = 0.042).
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: fiber_tertile_multivariate_pfs

#line 174-175
cox_pfs_gu_mv_tertile <-coxph(Surv(pfs_mo,
                               pod_status)~  fiber_tertile_factor + 
                            strata(age_cat) + 
                            ecog_1to5_final +
                            cancer_RCCvsUC_NAME +
                            stage.simplified.factor,
                          gu_cohort_data)

variable = "fiber_tertile_factorT3 (highest)"
sum_pfs_gu_tertile_mv <- summary(cox_pfs_gu_mv_tertile)
cat(paste(
  "-----------------------------------------------\n\np value fiber highest vs lowest tertile multivariate, PFS:",
  round(sum_pfs_gu_tertile_mv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_gu_tertile_mv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_gu_tertile_mv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_gu_tertile_mv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

```
We identified that each log-fold increase in fecal butyrate-production gene abundance was associated with a decrease in overall mortality, HR 0.44 (95% CI, 0.27-0.73; p =0.002) in a multivariable Cox model adjusted for age, performance status, BMI, smoking status and stratified on diagnosis.
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: total_buty_multivariate_os

#line 204-206
cox_os_mst_total_buty_mv <-coxph(Surv(tt_os_d, event_os)~ log(total) + 
                                age10 + 
                                strata(cohort) + 
                                imputed_ecog + #imputed to modal value
                                bmi + 
                                imputed_smoking_status, #imputed to modal value
                          mixed_solid_tumor)

variable = "log(total)"
sum_os_mst_total_buty_mv <- summary(cox_os_mst_total_buty_mv)
cat(paste(
  "-----------------------------------------------\n\np value log(total butyrate producing genes), OS:",
  round(sum_os_mst_total_buty_mv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_os_mst_total_buty_mv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_os_mst_total_buty_mv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_os_mst_total_buty_mv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))
```
We observe a comparable HR of 0.6 for PFS (95% CI, 0.39-0.96; p = 0.03).
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: total_buty_multivariate_pfs

#line 207-208
cox_pfs_mst_total_buty_mv <-coxph(Surv(tt_pfs_d, pfs_event)~ log(total) + 
                                age10 + 
                                strata(cohort) + 
                                imputed_ecog + #imputed to modal value
                                bmi + 
                                imputed_smoking_status, #imputed to modal value
                          mixed_solid_tumor)

variable = "log(total)"
sum_pfs_mst_total_buty_mv <- summary(cox_pfs_mst_total_buty_mv)
cat(paste(
  "-----------------------------------------------\n\np value log(total butyrate producing genes), PFS:",
  round(sum_pfs_mst_total_buty_mv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_mst_total_buty_mv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_mst_total_buty_mv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_mst_total_buty_mv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))
```
We then subdivided the butyrate-producing genes according to their annotation as steps in four substrate-specific fermentation subpathways as described in Vital et al.19 (acetyl-CoA [ACoA], glutarate, succinate, lysine) and investigated associations with OS for each, applying a Bonferroni correction for multiple testing (significance at p < 0.0125). The ACoA fermentation pathway, which is the fermentation pathway involved in deriving butyrate from fiber, accounted for the association between butyrate-production capacity, as this was the only sub-pathway for which a statistically significant difference was observed in OS (Bonferroni-adjusted Wald p = 0.002 ...
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: acoa_vs_taxa_os

#line 211 ish
model1_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(rum)))
model2_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor%>% mutate(test_v = log(rum_lach)))
model3_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridiales)))
model4_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridia)))
model5_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(eval(faecalibacterium_prausnitzii+1e-4))))
model6_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(pyruvate)))

models_mixed_solid_tumor_os <- list(model6_msk_os, model1_msk_os, model2_msk_os, model3_msk_os, model4_msk_os, model5_msk_os)

results_df <- data.frame()

for (i in seq_along(models_mixed_solid_tumor_os)) {
  model <- models_mixed_solid_tumor_os[[i]]
  summary_model <- summary(model)


  hr <- round(summary_model$coefficients["test_v", "exp(coef)"], digits = 2)
  lower_ci <- round(summary_model$conf.int["test_v", "lower .95"], digits = 2)
  upper_ci <- round(summary_model$conf.int["test_v", "upper .95"], digits = 2)
  pvalue <- round(summary_model$coefficients["test_v", "Pr(>|z|)"], digits = 4)


  results_df <- rbind(results_df, data.frame(
    Model = paste0("model", i),
    HR = hr,
    Lower_CI = lower_ci,
    Upper_CI = upper_ci,
    pvalue = pvalue
  ))
}
p_adjusted <- p.adjust(results_df$pvalue, method = "bonferroni")
#First p value is the one for ACoA
cat(paste(
  "-----------------------------------------------\n\nBonferroni Adju p value log(ACoA), OS:",
  p_adjusted[1],
  "\n\n-----------------------------------------------\n\n"
))

```
A significant association with PFS was also observed (Wald p = 0.015, Bonferroni adj. Wald p = 0.090).
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: acoa_vs_taxa_pfs

#line 211 ish
model1_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(rum)))
model2_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor%>% mutate(test_v = log(rum_lach)))
model3_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridiales)))
model4_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridia)))
model5_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(eval(faecalibacterium_prausnitzii+1e-4))))
model6_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(pyruvate)))


models_mixed_solid_tumor_pfs <- list(model6_msk_pfs, model1_msk_pfs, model2_msk_pfs, model3_msk_pfs, model4_msk_pfs, model5_msk_pfs)

results_df <- data.frame()


for (i in seq_along(models_mixed_solid_tumor_pfs)) {
  model <- models_mixed_solid_tumor_pfs[[i]]
  summary_model <- summary(model)


  hr <- round(summary_model$coefficients["test_v", "exp(coef)"], digits=2)
  lower_ci <- round(summary_model$conf.int["test_v", "lower .95"], digits=2)
  upper_ci <- round(summary_model$conf.int["test_v", "upper .95"], digits=2)
  pvalue <- round(summary_model$coefficients["test_v", "Pr(>|z|)"], digits=3)


  results_df <- rbind(results_df, data.frame(
    Model = paste0("model", i),
    HR = hr,
    Lower_CI = lower_ci,
    Upper_CI = upper_ci,
    pvalue = pvalue
  ))
}

p_adjusted <- p.adjust(results_df$pvalue, method = "bonferroni")
#First p value is the one for ACoA
cat(paste(
  "-----------------------------------------------\n\n",
  "Bonferroni Adju p value log(ACoA), PFS:",
  p_adjusted[1],
  "\n\n-----------------------------------------------\n\n",
  
  "Not Adj p value log(ACoA), PFS:",
  results_df$pvalue[1],
  "\n\n-----------------------------------------------\n\n"
))

```
Patients in this cohort were randomly assigned to receive a live bacterial supplement or placebo throughout treatment, and in a multivariable model including log-transformed ACoA-pathway gene abundance and treatment assignment, increased gene abundance was associated with significant improvement in PFS (HR 0.29 [0.11, 0.75], p=0.011)
```{r}
#| warning: false
#| echo: true
#| include: true
#| label: external_rcc_aca_pfs

#line 241

cox_pfs_ercc_log_acoa_mv <-coxph(Surv(pfsd, progressed_per_protocol)~ log(pyruvate)+cbm588, 
                          external_rcc %>% 
                            filter(!is.na(progressed_per_protocol)) %>%
                            mutate(
                              pfsd = as.numeric(as.Date(date_of_progression_or_last_f_u) - as.Date(x1st_i_n)) 
                              )
)

variable = "log(pyruvate)"
sum_pfs_ercc_log_acoa_mv <- summary(cox_pfs_ercc_log_acoa_mv)
cat(paste(
  "-----------------------------------------------\n\np value log(acoa), external RCC PFS:",
  round(sum_pfs_ercc_log_acoa_mv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_pfs_ercc_log_acoa_mv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_pfs_ercc_log_acoa_mv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_pfs_ercc_log_acoa_mv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

```

```{r}
#| warning: false
#| echo: true
#| include: true
#| label: external_melanoma_pfs_log_acoa

#line 250 ish

cox_em_pfs_mv <- coxph(Surv(pfs_d, pfsevent)~log(pyruvate) + age10 + adv_substage, external_melanoma)


variable = "log(pyruvate)"
sum_em_pfs_mv <- summary(cox_em_pfs_mv)
cat(paste(
  "-----------------------------------------------\n\np value log(acoa), external RCC PFS:",
  round(sum_em_pfs_mv$coefficients[variable, "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n",
  "HR:",
  round(sum_em_pfs_mv$coefficients[variable, "exp(coef)"],digits = 2),
  "[95% CI:",
  round(sum_em_pfs_mv$conf.int[variable, "lower .95"],digits = 2),
  round(sum_em_pfs_mv$conf.int[variable, "upper .95"],digits = 3),
  "]"
))

```


