--- 
title: |
  Extended Figure 1: PFS vs Fiber by Disease Type
---



```{r}
#| echo: false
#| warning: false
#| include: false
#| label: library_data_imports

library(dplyr)
library(survival)
library(survminer)
library(forestplot)
source("utility_funcs.R")

gu_cohort_data <- readRDS("../data/gu_survey_cohort.RDS")

makeSpline <- function(df, survtime, survevent, var,
                         knots=knots, centerpoint = NULL,
                         title=paste("Spline with", knots, "knots"),
                         xlab = xlab, ylab=ylab,
                         xticks = NULL) {
    
    require(survival)
    require(tidyverse)
    
    model_spline <- coxph(Surv(survtime, survevent)~pspline(var,knots), df) #
    ptemp <- termplot(model_spline, se=T, plot=F)
    
    splinem <- ptemp$var
    
    if(is.null(centerpoint)){centerpoint = median(var, na.rm=T)}
   
    if(centerpoint %in% splinem$x){
      c_row = which(splinem$x == centerpoint)
    } else {
      c_row = which(abs(splinem$x - centerpoint) == min(abs(splinem$x - centerpoint)))
    }
    
    center <- splinem$y[c_row]
    
    ytemp <- splinem$y + outer(splinem$se, c(0, -1.96, 1.96), '*')
    
    exp_ytemp <- exp(ytemp - center)
    
    # Create a spline_data frame for plotting
    spline_data <- data.frame(xvar = splinem$x, Estimate = exp_ytemp[,1],
                              Lower = exp_ytemp[,2], Upper = exp_ytemp[,3])
    
    ggplot(spline_data, aes(x = xvar)) +
      geom_hline(yintercept=1, lty=2)+
      geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "grey80", alpha = 0.5) +
      geom_line(aes(y = Estimate), color = "blue") +
      geom_rug(sides = "b") +  # Add rug plot at the bottom ('b') of the plot
      scale_y_log10() +  # Log scale for y-axis
      scale_x_log10() +
      labs(x = xlab, y = ylab, title=title) +
      theme_classic()
  }
```

The data we are visualizing here has the following structure:


```{r}
#| warning: false
#| include: false
#| label: load_data

gu_cohort_data <- readRDS("../data/gu_survey_cohort.RDS")

#pull out cohorts:

mUCdataFiber=filter(gu_cohort_data,cancer_type!=4)
dataUCavelumabFiber=filter(mUCdataFiber,io_regimen==5)
dataUC_NOTavelumabFiber=filter(mUCdataFiber,io_regimen!=5)

mRCdataFiber=filter(gu_cohort_data,cancer_type==4)
dataRCIpiNivoFiber=filter(mRCdataFiber,io_regimen==9 | io_regimen==17)
dataRCioTKIFiber=filter(mRCdataFiber,io_regimen==8 | io_regimen==18 | io_regimen==10 | io_regimen==11 | io_regimen==16 | io_regimen==12)

```

## Panel A

Mortality Hazard for PFS by daily average fiber intake, visualized via partial smoothing spline for pateints with Renal Cell Cancer.  Ticks along the x axis represent individual patients. 

```{r}
#| warning: false
#| include: true
#| label: panel_a

makeSpline(gu_cohort_data, gu_cohort_data$pfs_mo, gu_cohort_data$pod_status, 
           gu_cohort_data$gram_fiber_per_Kcal,
           xlab = "(Grams of fiber)/(1,000 kilocalories)", ylab="Hazard ratio (95% CI)",
           title="Smoothing spline for progression-free survival",
           knots=3) +
  scale_x_continuous(breaks = c(5,10,15,20))

#ggsave("svg/extended_figure_1_part_a_fiber_density_spline.svg", w=8, h = 5)
```


## Panel B

Mortality Hazard for PFS by daily average fiber intake, visualized via partial smoothing spline for pateints with Urothelial Cancer.  Ticks along the x axis represent individual patients. 

```{r}
#| warning: false
#| include: true
#| label: panel_b


makeSpline(gu_cohort_data, gu_cohort_data$os_mo, gu_cohort_data$os_status, 
           gu_cohort_data$gram_fiber_per_Kcal,
           xlab = "(Grams of fiber)/(1,000 kilocalories)", ylab="Hazard ratio (95% CI)",
           title="Smoothing spline for overall survival",
           knots=3) +
  scale_x_continuous(breaks = c(5,10,15,20))
#ggsave("svg/extended_figure_1_part_b_spline_os.svg", w=8, h = 5)
```

## Panel C

KM Surv Plot above below 15g/1000kcal

```{r}
#| warning: false
#| include: true
#| label: panel_c
kmPFS <- survfit(Surv(pfs_mo, pod_status) ~ gram_fiber_per_Kcal_aboveRecommended,
                 data = gu_cohort_data)

ggsurvplot(kmPFS,risk.table=TRUE,
           xlab="Time (months)", 
           ylab="Progression-free survival",
           pval = TRUE,
           legend=c(0.9,0.8),
           title="Patients with urothelial or renal cell cancer on ICB",
           legend.title="Fiber intake",
           legend.labs=c("<15 g/1,000 kcal","15+ g/1,000 kcal"),
           conf.int=TRUE,
           break.x.by=6,
           axes.offset=FALSE,
           xlim=c(0,25),
           data = gu_cohort_data)


```

