--- 
title: |
  Figure 2 - Butyrate Production Capacity from ACoA Pathway and ICB outcome
format:
  html:
    css: styles.css
---



```{r}
#| echo: false
#| warning: false
#| include: false
#| label: library_data_imports

library(tidyverse)
library(ggplot2)
library(phyloseq)
library(microViz)
library(survival)
library(survminer)
library(viridis)
library(gtsummary)
library(yingtools2)
source("utility_funcs.R")


```


Here are the column names of the data we are plotting:

fouraminobutyrate, glutarate, lysine and pyruvate represent different parts of the butyrate synthesis pathway. 

```{r}
#| warning: false
#| echo: false
#| label: load_data
mixed_solid_tumor <- readRDS("../data/mixed_solid_tumor.RDS")
external_rcc <- readRDS("../data/external_rcc.rds")
external_melanoma_cohort <- readRDS("../data/external_melanoma_metadata.RDS")
```

### Panel A

Survival in Mixed Solid Tumor Cohort (n = 147) by ACoA pathway gene abundance in stool above/below the median.

First for overall survival (OS):

```{r}
#| warning: false
#| label: panel_a_os
#| echo: false
#| include: true

fit<- survfit(Surv(tt_os_months, event_os) ~ med_p, data = mixed_solid_tumor)
names(fit$strata) <- gsub("med_p=", "", names(fit$strata))
p <- ggsurvplot(fit,  size = 1,  # change line size
           linetype = "solid", # use solid line
           break.time.by = 12, # break time axis by 12 months
           palette = c("aquamarine4", "coral4"), # custom color palette
           conf.int = TRUE, # Add confidence interval
           pval = TRUE, # Add p-value,
           legend.title="",
           risk.table = TRUE, risk.table.y.text.col = TRUE,
           ylab = "Survival %",
           xlab = "Time (Months)",
           xlim = c(0,48)
           ) 
print(p)

```

Then looking at Progression Free Survival (PFS) instead of OS:

```{r}
#| warning: false
#| echo: false
#| include: true
#| label: panel_a_pfs


fit<- survfit(Surv(tt_pfs_months, pfs_event) ~ med_p, data = mixed_solid_tumor)
names(fit$strata) <- gsub("med_p=", "", names(fit$strata))
p <- ggsurvplot(fit,  size = 1,  # change line size
           linetype = "solid", # use solid line
           break.time.by = 12, # break time axis by 12 months
           palette = c("aquamarine4", "coral4"), # custom color palette
           conf.int = TRUE, # Add confidence interval
           pval = TRUE, # Add p-value,
           legend.title="",
           risk.table = TRUE, risk.table.y.text.col = TRUE,
           ylab = "Survival %",
           xlab = "Time (Months)",
           xlim = c(0,48)
           ) 
print(p)
```

### Panel B

 Restricted cubic spline for Observational cohort adjusted for age, diagnosis, and performance status identifies increasing hazard of overall mortality associated with below-median (<504 RPKM) ACoA pathway gene abundance among ICB recipients. Ticks along the x-axis reflect individual samples. 

```{r}
#| warning: false
#| echo: true
#| include: true
#| label: panel_b

cox_os_mv_spline <- coxph(Surv(tt_os_d, event_os) ~ rms::rcs(pyruvate,4) + age10 + factor(cohort) + ecog + bmi + imputed_smoking_status, mixed_solid_tumor)
summary(cox_os_mv_spline)
ptemp <- termplot(cox_os_mv_spline, se=T, plot=F)
buterm <- ptemp$pyruvate
center <- buterm$y[which(buterm$y == median(buterm$y))]
ytemp <- buterm$y + outer(buterm$se, c(0, -1.96, 1.96), '*')
exp_ytemp <- exp(ytemp - center)

spline_data <- data.frame(buty = buterm$x, Estimate = exp_ytemp[,1],
                          Lower = exp_ytemp[,2], Upper = exp_ytemp[,3])

ggplot(spline_data, aes(x = buty)) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "grey80", alpha = 0.5) +
  geom_line(aes(y = Estimate), color = "blue") +
  geom_rug(sides = "b") +  # Add rug plot at the bottom ('b') of the plot
  scale_y_log10() +  # Log scale for y-axis
  labs(
    title = "Spline for Overall Survival",
    x = "ACoA-pathway genes present (RPKM)",
    y = "Hazard ratio") +
  theme_minimal()
ggsave("svg/updated_fig_2_spline_with_bmi_smoking.svg", w = 3, h = 3) 

```

```{r}
#| warning: false
#| echo: true
#| include: true
#| label: panel_b_pval

#also fit model without spline
cox_os_mv <- coxph(Surv(tt_os_d, event_os) ~ log(pyruvate) + age10 + factor(cohort) + ecog + bmi + imputed_smoking_status, mixed_solid_tumor)


cat(paste(
  "-----------------------------------------------\n\nWald p value for ACoA pathway abundance, OS pyruvate models with spline:",
  round(summary(cox_os_mv_spline)$coefficients["rms::rcs(pyruvate, 4)pyruvate", "Pr(>|z|)"],digits = 5),
  "\n\n-----------------------------------------------\n\n"
))

cat(paste(
  "-----------------------------------------------\n\nWald p value for ACoA pathway abundance, OS pyruvate continous:",
  round(summary(cox_os_mv)$coefficients["log(pyruvate)", "Pr(>|z|)"],digits = 5),
  "\n\n-----------------------------------------------\n\n"
))
```


We will now plot some microbiome data, using a `phyloseq` object that contains our sequencing data.  This object contains a list of the taxonomy of organisms identified, the abundance in each sample, and any associated sample specific metadata in one object. 

Principal components analysis (PCA) of Observational cohort sample (n = 147, 1 per patient) taxonomic abundance grouped at the family level shows increasing ACoA-pathway gene abundance from left to right along the first principal component (x-axis). Each point represents a single fecal sample. PCA loading vectors (arrows) identify increasing Ruminococcaceae and Lachnospiraceae abundance in the direction corresponding to increased ACoA-pathway gene abundance, versus greater Bacteroidales abundance (grouped at the order level) in the direction of decreased ACoA-pathway gene presence. 

### Panel C



```{r}
#| warning: false
#| include: false
#| label: prepare_data_panel_c

# Load in our prepared deidentified ps:

ps <- readRDS("../public_data/sequencing_data/plotting_phyloseq_deidentified.rds")

sample_data(ps) <- sample_data(ps) %>% 
  as.matrix() %>%
  as.data.frame() %>% 
  mutate(SAMPID = as.character(SAMPID)) %>%
  rownames_to_column("sampl_names") %>%
  left_join(mixed_solid_tumor %>% mutate(SAMPID = as.character(SAMPID)) %>% select(c(log_acoa, SAMPID)), by = "SAMPID") %>%
  column_to_rownames("sampl_names") %>%
  sample_data()
```
First looking at how log ACOA distribution looks superimposed over the PCOA of Stool Samples:

```{r}
#| warning: false
#| label: panel_c_part_1
ps %>%
  ps_filter(SAMPID %in% mixed_solid_tumor$SAMPID) %>%
  tax_fix() %>%
  tax_transform(trans = "compositional", rank = "Species") %>%
  dist_calc("aitchison") %>%
  ord_calc("PCoA") %>%
  ord_plot(fill = "log_acoa", size = 4, shape=21, alpha=0.6,
           plot_taxa = c(1:5)) +
  scale_fill_gradientn(colours=c("#2b3f76", "#f0f0f0", "#b2182b"), values=c(0, 0.5, 1)) +
  theme_classic(base_size = 16) +  
  scale_x_reverse() + 
  labs(
    fill = "Log ACOA"
  )
```
Next looking at Ruminococcaeceae:

```{r}
#| warning: false
#| label: panel_c_part_2
rum_ps <- ps
sample_data(rum_ps)$log_rum = log10(as.numeric(sample_data(ps)$Ruminococcaceae) + 0.0001)
rum_ps %>%
  ps_filter(SAMPID %in% mixed_solid_tumor$SAMPID) %>%
  tax_fix() %>%
  tax_transform(trans = "compositional", rank = "Species") %>%
  dist_calc("aitchison") %>%
  ord_calc("PCoA") %>%
  ord_plot(fill = "log_rum", size = 4, shape=21, alpha=0.6,
           plot_taxa = c(1:5)) +
  scale_fill_gradientn(colours=c("#2b3f76", "#f0f0f0", "#b2182b"), values=c(0, 0.65, 1)) +
  theme_classic(base_size = 16) + 
  scale_x_reverse() + 
  labs(
    fill = "Log10 Rel Abund.\nRuminococcaceae"
  )
```
Lastly looking at Bacteroidaceae:

```{r}
#| warning: false
#| label: panel_c_part_3

bac_ps <- ps
sample_data(bac_ps)$log_bac = log10(as.numeric(sample_data(ps)$Bacteroidaceae) + 0.0001)
bac_ps %>%
  ps_filter(SAMPID %in% mixed_solid_tumor$SAMPID) %>%
  tax_fix() %>%
  tax_transform(trans = "compositional", rank = "Species") %>%
  dist_calc("aitchison") %>%
  ord_calc("PCoA") %>%
  ord_plot(fill = "log_bac", size = 4, shape=21, alpha=0.6,
           plot_taxa = c(1:5)) +
  scale_fill_gradientn(colours=c("#2b3f76", "#f0f0f0", "#b2182b"), values=c(0, 0.75, 1)) +
  theme_classic(base_size = 16) + 
  scale_x_reverse() + 
  labs(
    fill = "Log10 Rel Abund.\nBacteroidaceae"
  )
```

### Panel D

ACoA-pathway gene abundance was associated with response to ICB among patients with advanced renal cell cancer receiving ipilimumab plus nivolumab in the External Renal Cancer Cohort (n = 39). 

```{r}
#| warning: false
#| label: panel_d



#Figure 2D
ggplot(external_rcc, aes(x = response, y = pyruvate)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, linewidth = 0.6, color = "black") +
  geom_jitter(aes(fill = response), shape = 21, width = 0.15, size = 5, stroke = 0.3, color = "black", alpha = 0.7) +
  scale_fill_manual(values = c("#999999", "#E69F00")) +
  scale_x_discrete(labels = c("Non-responder", "Responder")) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "none",
    axis.line = element_line(color = "black", linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.5)
  ) +
  xlab("") +
  ylab("Acetyl-CoA pathway\ngene abundance (RPKM)")


wilcox.test(external_rcc$pyruvate ~ external_rcc$response)


group_n <- external_rcc %>% count(response)
print(group_n)

#adjusted for CMB588
summary(glm(response01 ~ log1p(pyruvate) + cbm588, data = external_rcc , family = binomial(link = "logit")))
```



### Panel E

Univariate Cox regression analyses PFS relationship to ACoA gene abundance in external cohort.

```{r}
#| warning: false
#| label: panel_e_external_data_pfs_mv


model1_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort %>% mutate(test_v = log(rum)))
model2_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort%>% mutate(test_v = log(rum_lach)))
model3_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort%>% mutate(test_v = log(clostridiales)))
model4_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort%>% mutate(test_v = log(clostridia)))
model5_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort%>% mutate(test_v = log(eval(faecalibacterium_prausnitzii+1e-7))))
model6_wargo_pfs <- coxph(Surv(pfs_d, pfsevent)~test_v + age10 + adv_substage, external_melanoma_cohort%>% mutate(test_v = log1p(pyruvate)))

models_external_melanoma_os <- list(model6_wargo_pfs, model1_wargo_pfs, model2_wargo_pfs, model3_wargo_pfs, model4_wargo_pfs, model5_wargo_pfs)
summary(model6_wargo_pfs)
results_df <- data.frame()

for (i in seq_along(models_external_melanoma_os)) {
  model <- models_external_melanoma_os[[i]]
  summary_model <- summary(model)

  hr <- round(summary_model$coefficients["test_v", "exp(coef)"], digits=2)
  lower_ci <- round(summary_model$conf.int["test_v", "lower .95"], digits=2)
  upper_ci <- round(summary_model$conf.int["test_v", "upper .95"], digits=2)
  pvalue <- round(summary_model$coefficients["test_v", "Pr(>|z|)"], digits=3)

  results_df <- rbind(results_df, data.frame(
    Model = paste0("model", i),
    HR = hr,
    Lower_CI = lower_ci,
    Upper_CI = upper_ci,
    pvalue = pvalue
  ))
}

results_df$source = rep("wargo", 6)
results_df$modelname = c("log(ACoA)", "log(Ruminococcaceae)", "log(Ruminococcacaeae + Lachnospiraceae)",
                             "log(Clostridiales)", "log(Clostridia)", "log(F prausnitizii + 1e-7)")

ggplot(results_df) +
  geom_point(aes(y=fct_rev(fct_reorder(modelname, HR)), x=HR)) +
  geom_errorbarh(aes(y=fct_rev(fct_reorder(modelname, HR)), xmin=Lower_CI, xmax=Upper_CI), height=0.5)+
  geom_vline(xintercept=1, lty=2, col="gray70") +
  scale_x_continuous(trans="log") +
  theme_classic() +
  labs(title = "HR (95% CI) multivariable Cox PFS\nExternal Melanoma Cohort",
       subtitle ="", y="Variable", x="HR")


```

<h3 class="nogo" >Panel E MSK Data</h3>
Univariate Cox regression analyses PFS relationship to ACoA gene abundance in MSK Observational Data.

```{r}
#| warning: false
#| label: panel_e_internal_data_pfs_mv

model1_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(rum)))
model2_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor%>% mutate(test_v = log(rum_lach)))
model3_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridiales)))
model4_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridia)))
model5_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(eval(faecalibacterium_prausnitzii+1e-4))))
model6_msk_pfs <- coxph(Surv(tt_pfs_d, pfs_event)~test_v+age10+strata(cohort)+imputed_ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log1p(pyruvate)))


models_mixed_solid_tumor_pfs <- list(model6_msk_pfs, model1_msk_pfs, model2_msk_pfs, model3_msk_pfs, model4_msk_pfs, model5_msk_pfs)

results_df <- data.frame()


for (i in seq_along(models_mixed_solid_tumor_pfs)) {
  model <- models_mixed_solid_tumor_pfs[[i]]
  summary_model <- summary(model)


  hr <- round(summary_model$coefficients["test_v", "exp(coef)"], digits=2)
  lower_ci <- round(summary_model$conf.int["test_v", "lower .95"], digits=2)
  upper_ci <- round(summary_model$conf.int["test_v", "upper .95"], digits=2)
  pvalue <- round(summary_model$coefficients["test_v", "Pr(>|z|)"], digits=3)


  results_df <- rbind(results_df, data.frame(
    Model = paste0("model", i),
    HR = hr,
    Lower_CI = lower_ci,
    Upper_CI = upper_ci,
    pvalue = pvalue
  ))
}

results_df$source = rep("msk", 6)
results_df$type = rep("pfs", 6)
results_df$modelname = c("log(ACoA)","log(Ruminococcaceae)", "log(Ruminococcacaeae + Lachnospiraceae)",
                             "log(Clostridiales)", "log(Clostridia)", "log(F prausnitizii + 1e-7)")


ggplot(results_df) +
  geom_point(aes(y=fct_rev(fct_reorder(modelname, HR)), x=HR)) +
  geom_errorbarh(aes(y=fct_rev(fct_reorder(modelname, HR)), xmin=Lower_CI, xmax=Upper_CI), height=0.5)+
  geom_vline(xintercept=1, lty=2, col="gray70") +
  scale_x_continuous(trans="log") +
  theme_classic() +
  labs(title = "HR (95% CI) multivariable Cox PFS\nMixed Solid Tumor Cohort",
       subtitle ="", y="Variable", x="HR")


```

```{r}
#| warning: false
#| label: panel_e_internal_data_os_mv

model1_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(rum)))
model2_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor%>% mutate(test_v = log(rum_lach)))
model3_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridiales)))
model4_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(clostridia)))
model5_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log(eval(faecalibacterium_prausnitzii+1e-4))))
model6_msk_os <- coxph(Surv(tt_os_d, event_os)~test_v+age10+strata(cohort)+ecog+ bmi + imputed_smoking_status, mixed_solid_tumor %>% mutate(test_v = log1p(pyruvate)))


models_mixed_solid_tumor_os <- list(model6_msk_os, model1_msk_os, model2_msk_os, model3_msk_os, model4_msk_os, model5_msk_os)

results_df <- data.frame()

for (i in seq_along(models_mixed_solid_tumor_os)) {
  model <- models_mixed_solid_tumor_os[[i]]
  summary_model <- summary(model)


  hr <- round(summary_model$coefficients["test_v", "exp(coef)"], digits = 2)
  lower_ci <- round(summary_model$conf.int["test_v", "lower .95"], digits = 2)
  upper_ci <- round(summary_model$conf.int["test_v", "upper .95"], digits = 2)
  pvalue <- round(summary_model$coefficients["test_v", "Pr(>|z|)"], digits = 3)


  results_df <- rbind(results_df, data.frame(
    Model = paste0("model", i),
    HR = hr,
    Lower_CI = lower_ci,
    Upper_CI = upper_ci,
    pvalue = pvalue
  ))
}

results_df$source = rep("msk", 6)
results_df$type = rep("os", 6)
results_df$modelname = c("log(ACoA)","log(Ruminococcaceae)", "log(Ruminococcacaeae + Lachnospiraceae)",
                             "log(Clostridiales)", "log(Clostridia)", "log(F prausnitizii + 1e-7)")


ggplot(results_df) +
  geom_point(aes(y=fct_rev(fct_reorder(modelname, HR)), x=HR)) +
  geom_errorbarh(aes(y=fct_rev(fct_reorder(modelname, HR)), xmin=Lower_CI, xmax=Upper_CI), height=0.5)+
  geom_vline(xintercept=1, lty=2, col="gray70") +
  scale_x_continuous(trans="log") +
  theme_classic() +
  labs(title = "HR (95% CI) multivariable Cox OS\nMixed Solid Tumor Cohort",
       subtitle ="", y="Variable", x="HR")


```
