--- 
title: |
  Figure 1 - Hazard Ratio Analysis of Fiber Intake
---

This figure is largely concerned with fiber analysis in the Genito-urinary survey cohort.


```{r}
#| echo: false
#| warning: false
#| include: false
#| label: library_data_imports

library(dplyr)
library(survival)
library(survminer)
library(forestplot)
source("utility_funcs.R")
library(Hmisc)
library(RColorBrewer)
library(phyloseq)

ggsave_workaround <- function(g){survminer:::.build_ggsurvplot(x = g,
                                                               surv.plot.height = NULL,
                                                               risk.table.height = NULL,
                                                               ncensor.plot.height = NULL)}

ylgnbu_palette <- brewer.pal(n = 8, name = "YlGnBu")
custom_palette <- c(ylgnbu_palette[5], ylgnbu_palette[6], ylgnbu_palette[8])
```

The data we are visualizing here has the following structure:

```{r}
#| warning: false
#| include: false
#| label: load_data


gu_cohort_data <- readRDS("../data/gu_survey_cohort.RDS")
mixed_solid_tumor <- readRDS("../data/mixed_solid_tumor_phyloseq.RDS") %>% sample_data() %>% as.matrix() %>% as.data.frame()
external_melanoma <- readRDS("../data/external_melanoma_metadata.RDS")
external_rcc <- readRDS("../data/external_rcc.rds")

#pull out cohorts:

mUCdataFiber=filter(gu_cohort_data,cancer_type!=4)
dataUCavelumabFiber=filter(mUCdataFiber,io_regimen==5)
dataUC_NOTavelumabFiber=filter(mUCdataFiber,io_regimen!=5)

mRCdataFiber=filter(gu_cohort_data,cancer_type==4)
dataRCIpiNivoFiber=filter(mRCdataFiber,io_regimen==9 | io_regimen==17)
dataRCioTKIFiber=filter(mRCdataFiber,io_regimen==8 | io_regimen==18 | io_regimen==10 | io_regimen==11 | io_regimen==16 | io_regimen==12)
```

## Panel A

We start with some summaries of the cohorts used here:
```{r}
#| warning: false
#| include: true
#| echo: false
#| label: panel_a

cat(
  paste(
    "Genito-Urinary Survey Cohort:", nrow(gu_cohort_data),
    "Mixed Solid Tumor Cohort:", nrow(mixed_solid_tumor),
    "External Melanoma Cohort:", nrow(external_melanoma),
    "External RCC Cohort:", nrow(external_rcc),
    sep = "\n"
  )
)

```
## Panel B

Mortality Hazard for PFS by daily average fiber intake, visualized via partial smoothing spline.  Ticks along the x axis represent individual patients. 

```{r}
#| warning: false
#| include: true
#| echo: false
#| label: panel_b

p <- makeSpline(gu_cohort_data, gu_cohort_data$pfs_mo, gu_cohort_data$pod_status, 
           gu_cohort_data$aofib,
           xlab = "Daily average fiber intake (g/day)", ylab="Hazard ratio (95% CI)",
           title="Spline for progression-free survival",
           knots=2,
           centerpoint = 15)
print(p)

```

```{r}
#| warning: false
#| include: true
#| label: panel_b_pval

cox_pfs_allMet_uv <-coxph(Surv(pfs_mo,
                               pod_status)~  fiber_5g_increment,
                          gu_cohort_data)

cat(paste(
  "-----------------------------------------------\n\np value fiber in 5g increments, PFS:",
  round(summary(cox_pfs_allMet_uv)$coefficients["fiber_5g_increment", "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n"
))
summary(cox_pfs_allMet_uv)

```

## Panel C

Mortality Hazard for OS by daily average fiber intake, visualized via partial smoothing spline.  Ticks along the x axis represent individual patients. 

```{r}
#| warning: false
#| include: true
#| echo: false
#| label: panel_c

p<- makeSpline(gu_cohort_data, gu_cohort_data$os_mo, gu_cohort_data$os_status, 
           gu_cohort_data$aofib,
           xlab = "Daily average fiber intake (g/day)", ylab="Hazard ratio (95% CI)",
           title="Smoothing spline for overall survival",
           knots=2, 
           centerpoint = 15)
print(p)
```

```{r}
#| warning: false
#| include: true
#| label: panel_c_pval

cox_os_allMet_uv <-coxph(Surv(os_mo,
                              os_status)~  fiber_5g_increment,
                         gu_cohort_data)

cat(paste(
  "-----------------------------------------------\n\np value fiber in 5g increments, OS:",
  round(summary(cox_os_allMet_uv)$coefficients["fiber_5g_increment", "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n"
))
summary(cox_os_allMet_uv)
```

## Panel D

Kaplan-Meier PFS by fiber intake tertiles.

```{r}
#| warning: false
#| include: true
#| echo: false
#| label: panel_d


kmPFS_tert<-survfit(Surv(pfs_mo, pod_status)~fiber_tertile_factor,
               type="kaplan-meier", data=gu_cohort_data)

pfs_plot <- ggsurvplot(kmPFS_tert,
           risk.table=T,
           xlab=element_blank(), 
           ylab=element_blank(),
           pval = T,
           palette = custom_palette,
           title=element_blank(),
           censor.shape="|",
           censor.size = 1,
           size = .5,
           #legend = "none",
           legend.labs=c("Low", "Mid", "High"), # an ordered list
           legend.title = "Fiber\nConsumption\nTertile",
           conf.int=F,
           break.x.by=6,
           axes.offset=FALSE,
           xlim=c(0,25),
           data = gu_cohort_data)
print(pfs_plot)

#pfs_plot$plot <- pfs_plot$plot +
#  theme(
#    axis.text.y = element_blank(),
#    axis.text.x = element_blank()
#    )
#pfs_plot_to_save <- ggsave_survival_workaround(pfs_plot)
#ggsave("figures/svg/pfs_tertitle_plot.svg", plot = pfs_plot_to_save,w = 1.8, h = 1.5)
```

```{r}
#| warning: false
#| include: true
#| label: panel_d_pval

cox_pfs_allMet_mv <-coxph(Surv(pfs_mo,
                               pod_status)~ fiber_tertile_factor + 
                            strata(age_cat) + 
                            ecog_1to5_final +
                            cancer_RCCvsUC_NAME +
                            stage.simplified.factor,
                          gu_cohort_data
)

cat(paste(
  "-----------------------------------------------\n\nWald p value top vs bottom tertile, PFS:",
  round(summary(cox_pfs_allMet_mv)$coefficients["fiber_tertile_factorT3 (highest)", "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n"
))
```

## Panel E

Kaplan-Meier OS by fiber tertile. 

```{r}
#| warning: false
#| include: true
#| echo: false
#| label: panel_e

kmOS_tert<-survfit(Surv(os_mo, os_status)~fiber_tertile_factor,
               type="kaplan-meier", data=gu_cohort_data)


os_plot <- ggsurvplot(kmOS_tert,
           risk.table=T,
           xlab=element_blank(), 
           ylab=element_blank(),
           size = .5,
           pval = T,
           palette = custom_palette,
           censor.shape="|",
           censor.size = 1,
           title=element_blank(),
           #legend = "none",
           legend.labs=c("Low", "Mid", "High"), # an ordered list
           legend.title = "Fiber\nConsumption\nTertile",
           conf.int=FALSE,
           break.x.by=6,
           axes.offset=FALSE,
           xlim=c(0,25),
           data = gu_cohort_data) 
print(os_plot)

#os_plot$plot <- os_plot$plot +
#  theme(
#    axis.text.y = element_blank(),
#    axis.text.x = element_blank()
#    )
#
#
#os_plot_to_save <- ggsave_survival_workaround(os_plot) 
#ggsave("figures/svg/os_tertitle_plot.svg", plot = os_plot_to_save, w = 1.8, h = 1.5)
```

```{r}
#| warning: false
#| include: true
#| label: panel_e_pval

cox_os_allMet_mv <-coxph(Surv(os_mo,
                               os_status)~ fiber_tertile_factor + 
                            strata(age_cat) + 
                            ecog_1to5_final +
                            cancer_RCCvsUC_NAME +
                            stage.simplified.factor,
                          gu_cohort_data
)

cat(paste(
  "-----------------------------------------------\n\nWald p value top vs bottom tertile, PFS:",
  round(summary(cox_os_allMet_mv)$coefficients["fiber_tertile_factorT3 (highest)", "Pr(>|z|)"],digits = 3),
  "\n\n-----------------------------------------------\n\n"
))
```

## Panel F
Univariate HRs for PFS among clinically relevant subgroups in the highest versus lowest tertile of fiber consumption.

```{r}
#| warning: false
#| include: true
#| echo: true
#| label: panel_f

# Prep cohorts:
mUCdataFiber=filter(gu_cohort_data,cancer_type!=4)
dataUCavelumabFiber=filter(mUCdataFiber,io_regimen==5)
dataUC_NOTavelumabFiber=filter(mUCdataFiber,io_regimen!=5)

mRCdataFiber=filter(gu_cohort_data,cancer_type==4)
dataRCIpiNivoFiber=filter(mRCdataFiber,io_regimen==9 | io_regimen==17)
dataRCioTKIFiber=filter(mRCdataFiber,io_regimen==8 | io_regimen==18 | io_regimen==10 | io_regimen==11 | io_regimen==16 | io_regimen==12)

# Get pvalues:

cox_pfs_all_uv <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                       gu_cohort_data)
summary(cox_pfs_all_uv)

cox_pfs_RC <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                   mRCdataFiber)
summary(cox_pfs_RC)

cox_pfs_RC_IpiNivo <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                           dataRCIpiNivoFiber)
summary(cox_pfs_RC_IpiNivo)

cox_pfs_RC_TKI_IO <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                          dataRCioTKIFiber)
summary(cox_pfs_RC_TKI_IO)

cox_pfs_UC <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                   mUCdataFiber)
summary(cox_pfs_UC)

cox_pfs_UC_NOTavelumab <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                               dataUC_NOTavelumabFiber)
summary(cox_pfs_UC_NOTavelumab)

cox_pfs_UC_avelumab <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor,
                            dataUCavelumabFiber)
summary(cox_pfs_UC_avelumab)

forestplot_data <- tibble::tibble(mean  = c(0.4304,0.3603, 0.4570, 0.3079, 0.5800, 0.8845, 0.4854),
                                  lower = c(0.2223, 0.1444, 0.09054, 0.07621, 0.2224, 0.2940, 0.06718),
                                  upper = c(0.833, 0.8992, 2.307, 1.244, 1.513, 2.661, 3.507),
                                  study = c("All","RCC", "RCC ipi-nivo", "RCC TKI-ICB",
                                            "UC", "UC non-maintenance ICB", "UC maintenance ICB"),
                                  n = c("86","48","15","24","38","24","14"),
                                  events = c("55","28","8","12","27","19","8"),
                                  HR = c("0.43","0.36", "0.46", "0.31", "0.58", "0.88", "0.49"))

forestplot_data |>
  forestplot(labeltext = c(study, n, events, HR),
             clip = c(0.09, 2.5),
             xlog = TRUE,
             xlab = "Hazard ratio for PFS (95% CI)",
             xticks = c(0.1, 0.25, 0.5, 1.0, 2.5),
             title = "Univariable HRs for PFS comparing highest vs. lowest tertile of fiber intake",
             txt_gp = fpTxtGp(xlab = grid::gpar(fontsize = 25),
                              ticks = grid::gpar(fontsize = 25))) |>
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue") |> 
  fp_add_header(study = c("Subgroup"),
                n = c("n"),
                events = c("Events"),
                HR = c("HR"))

```

## Panel G

Multivariate HR for PFS adjusted for age, diagnosis, and performance status.

```{r}
#| warning: false
#| include: true
#| echo: true
#| label: panel_g

cox_pfs_all <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor + strata(age_cat) + 
                      ecog_1to5_final + cancer_RCCvsUC_NAME + stage.simplified.factor,
                    gu_cohort_data)
summary(cox_pfs_all)

cox_pfs_RC <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor + ageby10 + ecog_1to5_final,
                   mRCdataFiber) #could not adjust for stage because all patients with RCC were stage IV
summary(cox_pfs_RC)

cox_pfs_UC <-coxph(Surv(pfs_mo,pod_status)~ fiber_tertile_factor + strata(age_cat) + ecog_1to5_final + stage.simplified.factor,
                   mUCdataFiber)
summary(cox_pfs_UC)

forestplot_data <- tibble::tibble(mean  = c(0.4936, 0.4334, 0.70086),
                                  lower = c(0.2503, 0.16228, 0.2609),
                                  upper = c(0.9733, 1.1572, 1.883),
                                  study = c("All", "RCC", "UC"),
                                  n = c("86","48","38"),
                                  events = c("55","28","27"),
                                  HR = c("0.49","0.43","0.70"))

forestplot_data |>
  forestplot(labeltext = c(study, n, events, HR),
             clip = c(0.09, 2.5),
             xlog = TRUE,
             xlab = "Hazard ratio for PFS (95% CI)",
             xticks = c(0.1,0.25,0.5,1.0,2.5),
             title = "Multivariable HRs for PFS comparing highest vs. lowest tertile of fiber intake",
             txt_gp = fpTxtGp(xlab = grid::gpar(fontsize = 25),
                              ticks = grid::gpar(fontsize = 25))) |>
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue") |> 
  fp_add_header(study = c("Subgroup"),
                n = c("n"),
                events = c("Events"),
                HR = c("HR"))


```